---
title: 'R Guided Project 02: Creating An Efficient Data Analysis Workflow, Part 2'
output: html_document
---

```{r}
library(tidyverse)
library(lubridate)

sales <- read.csv("sales2019.csv")
```

This R Markdown file can be considered as a continuation of the previous analysis.  
  
The company has provided us more data on some of its 2019 book sales, and it wants us to extract some usable knowledge from it. It launched a new program encouraging customers to buy more books on July 1st, 2019, and it wants to know if this new program was successful at increasing sales and improving review quality.

# Data Exploration

```{r}
# How big is the dataset?
dim(sales)
```

```{r}
# What are the column names?
colnames(sales)

# What data is stored in the first 6 rows of the dataset?
head(sales)
```

The `date` column shows the date when the order of the books was made.  
`user_submitted_review` contains short reviews of particular titles.  
`title` contains titles of the books.  
`total_purchased` indicates the numbers of the books sold to that particular customer on a given day.  
`customer_type` determines whether the customer was an individual or a business.

```{r}
# What are the types of each of the columns?
str(sales)
```

The `total_purchased` column stores integer values. All other columns are of character data type.

```{r}
# How many missing values are there in particular columns?
colSums(is.na(sales))
```

Three columns of the dataset do not have any missing values, but the other two have hundreds of them.


# Handling Missing Data

```{r}
# Remove the rows with no 'user_submitted_review'
complete_sales <- sales %>% 
  filter(
    !is.na(user_submitted_review)
  )

# Make sure that the 'user_submitted_review' column does not contain any missing values
colSums(is.na(complete_sales))

# Calculate the mean of the 'total_purchased' column, without the missing values
purchase_mean <- complete_sales %>% 
  filter(!is.na(total_purchased)) %>% 
  pull(total_purchased) %>% 
  mean %>% print

# Assign this mean to all of the rows where 'total_purchased' was NA
complete_sales <- complete_sales %>% 
  mutate(
    total_purchased = if_else(is.na(total_purchased), 
                                as.integer(purchase_mean),
                                total_purchased)
  )

# Make sure that the 'total_purchased' column does not have missing values anymore
complete_sales %>% pull(total_purchased) %>% is.na %>% sum

# Check the first and last 6 rows of the data
head(complete_sales)
tail(complete_sales)
```


# Processing Review Data

```{r}
# Pull all unique values from the 'user_submitted_review' column
complete_sales %>% pull(user_submitted_review) %>% unique
```

The reviews range from outright hate ("Hated it") to positive ("Awesome!").  
The chunk of code below helps to determine which reviews are positive and which are negative.

```{r}
is_positive <- function(review) {
  review_positive = case_when(
    str_detect(review, "Awesome") ~ TRUE,
    str_detect(review, "OK") ~ TRUE,
    str_detect(review, "Never") ~ TRUE,
    str_detect(review, "a lot") ~ TRUE,
    TRUE ~ FALSE # The review did not contain any of the above phrases
  )
}
complete_sales <- complete_sales %>% 
  mutate(
    is_positive = unlist(map(user_submitted_review, is_positive))
  )

# Check the first and last 6 rows of the data
head(complete_sales)
tail(complete_sales)
```


# Comparing Book Sales Between Pre- and Post-Program Sales

```{r}
complete_sales <- complete_sales %>% 
  mutate(
    date_status = if_else(mdy(date) < ymd("2019/07/01"), "Pre", "Post")
  )
complete_sales %>% 
  group_by(date_status) %>% 
  summarize(
    books_purchased = sum(total_purchased)
  )
```

It doesn't seem that the program has increased sales. Maybe there were certain books that increased in sales?

```{r}
complete_sales %>% 
  group_by(date_status, title) %>% 
  summarize(
    books_purchased = sum(total_purchased)
  ) %>% 
  arrange(title, date_status)
```

It turns out that certain books actually got more popular after the program started - 'R For Dummies' and 'Secrets of R For Advanced Students' are those which increased their sales.


# Comparing Book Sales Within Customer Type

```{r}
complete_sales %>% 
  group_by(date_status, customer_type) %>% 
  summarize(
    books_purchased = sum(total_purchased)
  ) %>% 
  arrange(customer_type, date_status)
```

Based on the table, it looks like businesses started purchasing more books after the program! There was actually a drop in individual sales.


# Comparing Review Sentiment Between Pre- and Post-Program Sales

```{r}
complete_sales %>% 
  group_by(date_status) %>% 
  summarize(
    num_positive_reviews = sum(is_positive)
  )
```

There's slightly more reviews after the program, but this difference seems negligible.
